---
title: "homework 8"
author: "Anais Aoki"
date: '2023-02-27'
output: html_document
---

```{r setup, include=FALSE}

library(ggplot2)
library(tidyverse)
library(dplyr)
library(car)

library(nlme)

hungry_cat <- read.csv("blocked_experiment.csv")
limp <- read.csv("Limpet_grazing.csv")

```
plant = plant id number
block = block(13, lettered a through m)
AMF = fungi present, y or no (this is an IV)

```{r}

head(cat$cat_mass)
plot1 <- ggplot(hungry_cat, aes(x=cat_mass, y=damage)) +
  geom_boxplot() +
  facet_wrap(~AMF) +
  theme_bw()
plot1

model1 <- lm(cat_mass ~ damage * AMF, data = hungry_cat)
autoplot(model1)

#plot cat mass against block
plot2 <- ggplot(hungry_cat, aes(x=cat_mass, y=block)) +
  geom_boxplot() +
  facet_wrap(~AMF) +
  theme_bw()
plot2
```

#blocking example
```{r}
blockmodel <- lme(cat_mass ~ AMF * damage, random = ~1|block, data=hungry_cat, 
method = "ML")

#check residuals of the block model
plot(blockmodel) #checks for homoscedascity 
qqnorm(resid(blockmodel)) #check for normality of residuals
hist(resid(blockmodel)) #scores the data so that average is centered around zero and you have standard deviations on either size 


#Evaluate the fixed factors in some way:
options(contrasts = c("contr.sum","contr.poly")) #THEN REFIT BLOCKMODEL
blockmodel <- lme(cat_mass ~ AMF * damage, random = ~1|block, data=hungry_cat, 
method = "ML")
Anova(blockmodel, type=3) #type III ANOVA test -> looking at interaction
options(contrasts = c("contr.treatment","contr.poly")) #reset contrasts

#Or "manually" test fixed factors:
blockmodel2 <- update(blockmodel, ~.-AMF:damage) #shorthand to drop interaction
anova(blockmodel,blockmodel2) #likelihood ratio test
```

##Homework 8##
• Site is a blocking factor (1-8, convert to factor)
• LimpetTreat is a categorical predictor (IV), type of limpet grazing on plates
• Fo is the response variable (DV), amount of microalgae on plates
• TempC is a continuous covariate (IV), average temperature of plates
```{r}

limpet <- lme(Fo ~ LimpetTreat * TempC, random = ~1|as.factor(Site), data=limp, 
method = "ML")

plot(limpet)
qqnorm(resid(limpet)) #check for normality of residuals
hist(resid(limpet)) 
Anova(limpet, type = 3)


# diagnostic plots #
#Normal q-q plot checks for homoscedascity of residuals - it does not look terrible but does not look great either. There is a little variation in the distribution of residuals across the regression line (line is not straight and outliers of residuals are present at the lower and upper ends) 
#Histogram distribution of data is not even (std -4 and more data present in the upper limit of the mean (0))
#There is a significant difference in Fo (amount of microalgae on plates) between limpet treatment (p=5.814e-08, df =3). Temperature does not significantly differ in amount of microalge (0.06009, df=1), only when interaction of limpet treatment is present (p=6.285e-06, df =3)


```

#3/1 In class code
```{r}

lacebugs <- read.csv("lacebugs.csv")
head(lacebugs)

lacemod1 <- lme(lacebugs_m2 ~ treatment * light + time, random 
= ~1 | tree, data = lacebugs, na.action=na.omit)

plot(lacemod1)
qqnorm(resid(lacemod1)) #wedge shaped residuals need to log transform 
hist(resid(lacemod1)) 

#logd-transformed
lacemod2 <- lme(log(lacebugs_m2+1) ~ treatment * light + time, random = 
~1 | tree, data = lacebugs, method="ML")
plot(lacemod2)
qqnorm(resid(lacemod2)) #wedge shaped residuals need to log transform 
hist(resid(lacemod2)) 
```