---
title: "Aoki BIO606 Lec01"
output: html_document
date: '2023-01-18'
---
```{r, setup, include=FALSE}
library(tidyverse)
library(readr)
library(ggplot2)
library(dplyr)

snowmelt <- read.csv("NEE_snowmelt_Biol606.csv", header=TRUE)

```

1. Calculate mean and standard deviation of cumulative June-August C uptake ("CumC_uptake_Jun_Aug")
```{r}
mean(na.omit(CumC_uptake_Jun_Aug)) 
sd(na.omit(CumC_uptake_Jun_Aug))
```
2. Calculate the length of the snow free period using the day the last snow melted (SNOW_MELT_ DATE_MODIS) and the day of the first fall snowfall (SNOW_FALL_START_MODIS)
```{r}
snowfree <- (snowmelt$SNOW_FALL_START_MODIS-snowmelt$SNOW_MELT_DATE_MODIS)
snowfree
```

3.log transform precipitation_Jun_Aug and create a dataframe that includes the non-log transformed and the log-transformed precipitation variables, and the length of the snow free period. Save it as a .csv file.
```{r}
dfnew <- data.frame(nonlog_precip = (snowmelt$precipitation_Jun_Aug),
                    log_precip = (log(snowmelt$precipitation_Jun_Aug)),
                    freesnow = (snowmelt$SNOW_FALL_START_MODIS-snowmelt$SNOW_MELT_DATE_MODIS))

write.csv(dfnew, file="dfnew.csv")
```
4. Look at the file you created and check what it looks like. Note that the commands for mean and standard deviation produce a single number, but log-transforming creates a vector of numbers that can be stored as a new variable.
```{r}
dfnew <- read.csv("dfnew.csv", row.names = FALSE)
View(dfnew)
```
5. Look at the hypotheses below. Make a simple plot illustrating H1 and another for H2. What kind of plot did you make?
H1. The length of the snow free period is mostly controlled by fall temperatures.
H2. An earlier snow melt will result in an increased uptake of carbon by Arctic tundra
```{r}
#Hypothesis 1
#snow free is dependent on fall temperatures
plot(snowfree ~ airT_Sept_mean, data = snowmelt, xlab= "Mean Air Temp in September", ylab= "Snow Free Period", col="blue")
#Hypothesis 2
plot(SNOW_MELT_DATE_MODIS ~ CumC_uptake_Jun_Aug, xlab = "Carbon Uptake", ylab = "Snow Melt")
```
6. Start a new code chunk for Monday's assigned reading (Beckerman et al. chapter 4), and use it to create the code in the reading, including your own annotations too.
```{r}

# clear the decks 
rm(list = ls())

# get the data
compensation <- read.csv('compensation.csv')

#check out the data 
glimpse(compensation) #shows columns, counts and data type

ggplot(compensation, aes(x = Root, y = Fruit)) +
geom_point() #simple scatterplot

##Scatter plot ##
ggplot(compensation, aes(x = Root, y = Fruit, colour = Grazing)) + #adds color to points which also adds legend
geom_point(size = 5) + #larger data points
xlab("Root Biomass") + #x axis label
ylab("Fruit Production") + #y axis label
theme_bw() #removes grey background 

ggplot(compensation, aes(x = Root, y = Fruit, shape = Grazing)) + #changes data points shape by variable + legend
geom_point(size = 5) + #larger data points
xlab("Root Biomass") + #x axis label
ylab("Fruit Production") + #y axis label
theme_bw() #removes grey background 

## Box plot ##
ggplot(compensation, aes(x = Grazing, y = Fruit)) +
geom_boxplot() + #simple box plot
geom_point(size = 4, colour = 'light grey', alpha = 0.5) + #adds light grey data points 
xlab("Grazing treatment") +
ylab("Fruit Production") +
theme_bw()

## Histogram ##
ggplot(compensation, aes(x = Fruit)) +
geom_histogram(bins = 15, binwidth = 15) +
theme_bw() +
xlab("Grazing treatment") +
ylab("Fruit Production") +
## ‘stat_bin()‘ using ‘bins = 30‘. Pick better value with
## ‘binwidth‘

ggplot(compensation, aes(x = Fruit)) +
geom_histogram(bins = 10) +
  theme_bw()
ggplot(compensation, aes(x = Fruit)) +
geom_histogram(binwidth = 15) +
facet_wrap(~Grazing) + #splits plots by grazing type
theme_bw()

## Saving Graphs ##
ggsave("ThatCoolHistogramOfFruit.png") #saves to working directory


```



